HUME: Human UCCA-Based Evaluation of Machine Translation

### 2016-03-18 paper/main.tex
edits
shorten section 3

This section summarises the manual annotation procedure used
to compute the HUME measure. 
We denote the source sentence as $s$ and the translation as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements as to the translation quality of each semantic unit of $s$ relative to $t$,
where units are defined according to the UCCA annotation.
We note that UCCA annotation is performed once for every source sentence,
irrespective of the number of its translations we wish
to evaluate\footnote{Instead of relying on the source,
the evaluation could be based on the UCCA structure for the \emph{reference}
translation.
More details in the discussion in \secref{src-vs-ref}.
% This would allow monolingual annotators to take part in the
% evaluation. On the downside, the translation would have to be aligned to the
% reference and more importantly, we would risk inaccurate judgements due to
% the naturally occurring differences between the source and its reference
% translations.
}.

\paragraph{UCCA Annotation.}
We begin by creating UCCA annotations for the source sentence, following the
UCCA guidelines\footnote{All UCCA-related resources can be found
  here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}.
Let $L$ be the set of possible UCCA labels.
Given a sentence $s = (w_1, \ldots,w_n)$, its UCCA annotation is a DAG $G=(V,E,\ell)$,
where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
The semantic units for a sentence $s$ according to $G$
is the set of yields of nodes in $V$. 
We note that the translation is not UCCA-annotated.

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=0.45\textwidth]{ucca-tree-mteval-v2.png}
    \end{center}
  \caption{\label{fig:hume_tree_v2}
     HUME annotation of an UCCA tree with a word aligned example translation shown below. 
Atomic units are labelled using traffic lights and structural nodes are marked A or B.}
\end{figure}

In Figure~\ref{fig:hume_tree_v2} we can see an example of a possible HUME
annotation, with the translation shown in English for ease of comprehension.
When evaluating ``to NYC'' the annotator looks at the translation and sees the
word ``stateside''. This word captures the whole phrase and so we mark the
structural node with an atomic label. Here we choose ``Orange'' but the
annotator could equally have decided that this translation is wrong and marked
it as ``Red''. The ability to mark structural nodes with atomic labels allows
the annotator to account for translations which only correspond at the phrase
level. Another feature highlighted in this example is that the root node is
``Adequate''. This might seem confusing, as the relationship between the scenes
has changed in the translation. However, an important feature of HUME is that
penalises errors where they occur. In HUME, the error is caused by a wrong
translation of the linker ``After'' which translates as ``by'' and this is
marked as ``Red''. The root node still has two scenes linked by one linker and
the relationship between child nodes is maintained and therefore it is marked
``Adequate''.
% Ondrej removes the following, he thinks that errors are not supposed to be
% propagated in HMEANT either:
%In HMEANT, it is not clear whether errors should be propegated up the SRL hierarchy.

@@

\paragraph{Translation Evaluation.}
The basic strategy for computing HUME is to traverse the semantic units
of the source sentence, which correspond to the different arguments and relations expressed
in the text, and to mark the extent to which they have been correctly translated.
HUME aggregates the judgements of the users into a composite score, 
which reflects the overall extent to which the semantic content of $s$ is preserved in $t$.

Human annotation of the semantic units requires first deciding whether
a unit is {\it structural}, i.e., has meaning-bearing sub-units,
or {\it atomic}. In most cases, atomic units
correspond to individual words, but they may also correspond to unanalyzable
multi-word expressions, i.e., expressions that do not have meaningful sub-units
(e.g., idioms like ``takes two to tango'').
When a multi-word unit is labeled as atomic, all its sub-units' annotations are discarded
from the evaluation.

@@

Structural units have sub-units (children in the UCCA graph), which may themselves
either be atomic or structural. The labels assigned to such nodes reflect the extent to which
the relation between the sub-units is preserved in $t$.
For instance, a Scene unit represents the relation between the
Scene elements, such as the predicate (main relation), participants and adverbials.
Structural units are labeled as ``Adequate'' or ``Bad'', meaning
that the relation between the sub-units went wrong.
We use a three label system with atomic units, as
opposed to a two label one in structural units,
as atomic units often have slight errors, in which cases we do want to provide partial credit.

@@

\begin{compactitem}
\item The components of the translation are ordered differently from the source,
  such that the relation between them changes. For instance, where
  the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
  the semantic unit corresponding to the Scene will be marked as Bad, while
  the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
\item A superfluous word or phrase has been inserted into
  the translation. For instance, if ``dog bites man'' is translated into ``dog viciously bites man''.
\item One of the sub-units is missing from the translation.
  For instance, if ``dog bites man'' is translated into ``dog bites''. \bh{Note that in this case we also 
  get an error in the atomic unit, assuming that ``man'' is a semantic unit in the source, but for the other
  two cases there is no atomic unit error.}
\end{compactitem}

We note that HUME labels reflect adequacy, rather than fluency judgements.
Specifically, annotators are instructed to
label a unit as Adequate if its translation is understandable and preserves
the meaning of the source unit, even if its fluency is impaired.

In non-configurational languages, the relationship between the
governing and the dependent units is often expressed using morphological
properties of the dependents instead of their ordering
(this holds especially for the verb and its modifiers,
less so for components of noun phrases). Errors in morphology
should be expressed on the atomic units, so HUME can behave slightly differently
on configurational vs. non-configurational languages.

@@

>>>>

This section summarises the manual annotation procedure used
to compute the HUME measure. 
We denote the source sentence as $s$ and the translation as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements as to the translation quality of each semantic unit of $s$ relative to $t$,
where units are defined according to the UCCA annotation.
Note that UCCA annotation is performed once for every source sentence,
irrespective of the number of its translations we wish
to evaluate\footnote{Instead of relying on the source,
the evaluation could be based on the UCCA structure for the \emph{reference}
translation.
More details in the discussion in \secref{src-vs-ref}.
% This would allow monolingual annotators to take part in the
% evaluation. On the downside, the translation would have to be aligned to the
% reference and more importantly, we would risk inaccurate judgements due to
% the naturally occurring differences between the source and its reference
% translations.
}.

\paragraph{UCCA Annotation.}
We begin by creating UCCA annotations for the source sentence, following the
UCCA guidelines\footnote{All UCCA-related resources can be found
  here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}.
Let $L$ be the set of possible UCCA labels.
Given a sentence $s = (w_1, \ldots,w_n)$, its UCCA annotation is a DAG $G=(V,E,\ell)$,
where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
The semantic units for a sentence $s$ according to $G$
is the set of yields of nodes in $V$. 
The translation is not UCCA-annotated.

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=0.45\textwidth]{ucca-tree-mteval-v2.png}
    \end{center}
  \caption{\label{fig:hume_tree_v2}
     HUME annotation of an UCCA tree with a word aligned example translation shown below. 
Atomic units are labelled using traffic lights and structural nodes are marked A or B.}
\end{figure}
% Ondrej removes the following, he thinks that errors are not supposed to be
% propagated in HMEANT either:
%In HMEANT, it is not clear whether errors should be propegated up the SRL hierarchy.

@@

\paragraph{Translation Evaluation.}
The basic strategy for annotating HUME is to traverse the semantic units
of the source sentence, which correspond to the arguments and relations expressed
in the text, and to mark the extent to which they have been correctly translated.
HUME aggregates the judgements of the users into a composite score, 
which reflects the overall extent to which the semantic content of $s$ is preserved in $t$.

Annotation of the semantic units requires first deciding whether
a unit is {\it structural}, i.e., has meaning-bearing sub-units,
or {\it atomic}. In most cases, atomic units
correspond to individual words, but they may also correspond to unanalyzable
multi-word expressions, i.e., expressions that do not have meaningful sub-units
(e.g., idioms like ``takes two to tango'').
When a multi-word unit is labeled as atomic, its sub-units' annotations are discarded
from the evaluation.

@@

Structural units have sub-units (children in the UCCA graph), which may themselves
either be atomic or structural. The labels assigned to such nodes reflect the extent to which
the relation between the sub-units is preserved in $t$.
\XXX{(BH: don't follow this) For instance, a Scene unit represents the relation between the
Scene elements, such as the predicate (main relation), participants and adverbials.}
Structural units are labeled as ``Adequate'' or ``Bad'', meaning
that the relation between the sub-units went wrong.
We use a three label system with atomic units, as
opposed to a two label one in structural units,
as atomic units often have slight errors, in which cases we do want to provide partial credit.

@@

\begin{compactitem}
\item The components of the translation are ordered differently from the source,
  such that the relation between them changes. For instance, if
  the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
  the semantic unit corresponding to the Scene will be marked as Bad, while
  the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
\item A superfluous word or phrase has been inserted into
  the translation. For instance, if ``dog bites man'' is translated into ``dog viciously bites man''.
\item One of the sub-units is missing from the translation.
  For instance, if ``dog bites man'' is translated into ``dog bites''. \bh{Note that in this case we also 
  get an error in the atomic unit, assuming that ``man'' is a semantic unit in the source, but for the other
  two cases there is no atomic unit error.}
\end{compactitem}

HUME labels reflect adequacy, rather than fluency judgements.
Specifically, annotators are instructed to
label a unit as Adequate if its translation is understandable and preserves
the meaning of the source unit, even if its fluency is impaired.

In non-configurational languages, the relationship between the
governing and dependent units is often expressed using morphological
properties of the dependents instead of their ordering
(this holds especially for the verb and its modifiers,
less so for components of noun phrases). Errors in morphology
should be expressed on the atomic units, so HUME can behave differently
on configurational vs. non-configurational languages.

@@

In Figure~\ref{fig:hume_tree_v2} we can see an example of a possible HUME
annotation, with the translation shown in English for ease of comprehension.
When evaluating ``to NYC'' the annotator looks at the translation and sees the
word ``stateside''. This word captures the whole phrase and so we mark the
structural node with an atomic label. Here we choose ``Orange'' since it approximately captures the meaning
in this context. The ability to mark structural nodes with atomic labels allows
the annotator to account for translations which only correspond at the phrase
level. Another feature highlighted in this example is that the root node is
``Adequate''. This is because the root node still has two scenes linked by one linker and
so the relationship between child nodes is maintained. There is an error caused by a wrong
translation of the linker ``After'' which translates as ``by'' and this is
marked as ``Red'', but this error is marked at the point at which it occurs.
